Structured Versus Decorated Cospans

### 2022-04-18 cospans_compositionality_4.tex
I think I fixed all the problems, but there's a sentence Christina doesn't like yet.
One more tiny change - just making a sentence prettier.

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section \ref{Applications} with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use decorated cospans.

@@

We now show that the decorated cospan double category $D \lCsp$ of open dynamical systems is not isomorphic to a structured cospan double category via \cref{thm:equiv}. Recall that in that theorem we start with the data required to build a decorated cospan category, namely a symmetric lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat,\times)$, and show that if the resulting pseudofunctor $F \maps \A \to \SMC$ factors through $\Rex$, then the opfibration $U \maps \X = \inta F \to \A$ has a left adjoint $L \maps \A \to \X$. We then obtain an isomorphism between decorated and structured cospan double categories, $F \lCsp \cong {}_L \lCsp(\X)$. We now show that in the case at hand, where $F = D$ is the functor sending each finite set $S$ to the set of dynamical systems on $\R^S$, the opfibration $U$ does \emph{not} have a left adjoint. Thus, the conditions of \cref{thm:equiv} cannot hold in this case: $F$ does not factor through $\Rex$.

>>>>

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section \ref{Applications} with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, in some cases we need decorated cospans, for reasons we explain.

@@

We now show that the decorated cospan double category $D \lCsp$ of open dynamical systems is not isomorphic to a structured cospan double category via \cref{thm:equiv}. Recall that in this theorem we start with the data required to build a decorated cospan category, namely a symmetric lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat,\times)$, and show that if the resulting pseudofunctor $F \maps \A \to \SMC$ factors through $\Rex$, then the opfibration $U \maps \X = \inta F \to \A$ has a left adjoint $L \maps \A \to \X$. We then obtain an isomorphism between decorated and structured cospan double categories, $F \lCsp \cong {}_L \lCsp(\X)$. We now show that in the case at hand, where $F = D$ is the functor sending each finite set $S$ to the set of dynamical systems on $\R^S$, the opfibration $U$ does \emph{not} have a left adjoint. Thus, the conditions of \cref{thm:equiv} cannot hold in this case: $F$ does not factor through $\Rex$.
